Modelling the gas kinematics of an atypical Lyman alpha emitting compact dwarf galaxy

### 2017-07-13 paper/paper.tex
new appendix

\newcommand{\mg}[1]{\textcolor{blue}{\bf MG: \textit{#1}}}

@@

\mg{I would remove the next paragraph as this seems clear and not necessary to start with an ``apology''.}
We do not claim that our findings discard the existence of different
undiscovered models that could reproduce the observed features in \tol.
More involved models with multiple kinematic components can
in principle also be successful.
We focus on presenting two simple and successful models that
manage, for the first time, to account for \tol's \lya line main
features.

@@

Using the total number of degrees of freedom, $d.o.f.=104$ we divide
the results in two classes: low $\chi^2/d.o.f.<35$ and high
$\chi^2/d.o.f.>35$. \mg{Why was $35$ chosen as a boundary?}
We apply a random forest classification algorithm to find the
more relevant parameters in splitting the results into these two
classes. 
The algorithm uses $500$ trees for the classification and a
maximum of $3$ depth levels. 
To check for stability we repeat the computation $10$ times by
randomly subsampling the input data to use $80\%$ of the data as
the training set.

We also perform a Kolmogorov-Smirnov (K-S) test to compare each parameter
distribution in the best $25$ models against the parent distribution
of $2500$ models. 
If we obtain a p-value $<0.05$ for a given model parameter, we conclude that
this parameter does influence the $\chi^2$ result, as the distribution for
the best $\chi^2$ models is statistically different to the
distribution from the global sample of $2500$ models.

We finally compute the best values for the influential parameters as
the values that produce the minimum $\chi^2$.  
We estimate the 1-$\sigma$ uncertainty from a parabolic fit to the
$\chi^2$ as a function of the influential parameters around its
corresponding minimum.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.95\textwidth]{tree_result.pdf}
\caption{Classification tree example. 
The aim is to find the parameters that separate two
classes: low $\chi^2/d.o.f. < 35$ and high $\chi^2/d.o.f>35$. 
Each box describes the condition over the parameter of interest, the
level of sample impurity, the total number of samples and the value of the
number of samples in each class.  
In this example we randomly sample $80\%$ of the full data set of models to
start with $28$ models in the first class and
$1962$ models in the second class; the best way to increase the
probability to find a result with low $\chi^2/d.o.f.$ ($13$ models,
fourth bottom box, from left to the right)
is having the clump radial velocity, $v_{\infty,{\rm cl}}<157.0$ \kms;
the clump velocity dispersion, $\sigma_{{\rm cl}}>55.6$ \kms, and the
probability to have a \lya photon emitted in a clump, $P_{\rm cl}>0.683$.
The final classification is not perfect, showing that all other
parameters are also important but these three are the most relevant.  
\label{fig:tree}}
\end{center}
\end{figure*}

@@

With $14$ free parameters our first concern is discovering which are
the most relevant in reproducing \tol's line shape.
The results of the random forest classifier yield that the clump outflow velocity
$v_{\infty,\rm{cl}}$, the clump velocity dispersion $\sigma_{\rm cl}$ and the probability
that the \lya emission comes from the clumps $P_{\rm cl}$, are the
most influential parameters in finding a model with low
$\chi^2/d.o.f.$.

Figure \ref{fig:tree} shows as an illustration the results for a
single classification tree.
\mg{Sorry but it is not clear to me what this Figure shows \& how it was assembled. I suggest: leave the KS test in the main text as it is straightforward to understand, and move the random trees with more technical details / references about the implementations to the appendix -- and then point to it as a second probe from the main tetx. This way it would break the flow less.}
The tree starts with $28$ and $1962$ models in the low and high
$\chi^2/d.o.f.$ classes, respectively. 
We find that the best classification
yields $13$ and $44$ models in the low and high $\chi^2/d.o.f.$
classes, respectively, after selecting for $v_{\infty,{\rm cl}}<157.0$ \kms,
$\sigma_{{\rm cl}}>55.6$ \kms and $P_{\rm cl}>0.683$. 
The split in two classes is not perfect. 
This means that the 3 parameters in the
model are the most influential in finding models with low
$\chi^2/d.o.f.$, but that some combinations of the other 11 parameters
are also relevant to find a good fit.

From the K-S tests we confirm that the same 3 parameters are the
only ones that can be constrained by observations;
$v_{\infty,\rm{cl}}$ (p-value  $10^{-18}$), $\sigma_{\rm cl}$ (p-value
$10^{-4}$) $P_{\rm cl}$ (p-value $10^{-4}$).
Once again, this test does not mean the other parameters do not
influence the resulting spectra at all; it means that they cannot be
constrained from \tol's observations.

The best models are shown in Figure \ref{multiphaseresults}.
Left column shows the difference between the integrated distributions
of the full sample (2500 input models) and the models with the lowest
$\chi^2/d.o.f.$. 
The right column shows the actual $\chi^2/d.o.f.$ 
and its corresponding parameter value. 
Under these conditions we find $\sigma_{\rm cl}=$\sigmaclump,
$v_{\infty {\rm, cl}}=$\inftyclump\ and $P_{\rm cl}=$\probaclump.
\mg{These errors seem still pretty small to me. You obtained these via a Fisher matrix of a parabolic fit, right? How good was the fit? Are the off-diagonal elements included?}

Qualitatively this result can be understood as follows.
Due to the large fraction of \lya photons being emitted within the moving clumps ($P_{\rm cl}$), the `intrinsic' profile represents the kinematics of the clumps, that is, the width of the intrinsic line is set by $\sigma_{\mathrm{cl}}$ and its median offset mainly by $v_{\infty}$ \& $\beta_{\rm cl}$.
Furthermore, the relatively low mean number of clumps per line of sight ($f_{\mathrm{cl}}$) combined with the high velocity dispersion of the clumps implies the 
existence of low-density inter-clump
regions where \lya photons can freely propagate which leads to an emergent spectrum close to the intrinsic one as well as to the high flux at line center as observed.
Because the width of the observable spectrum is hence set primarily by $\sigma_{\rm cl}$, a lower velocity dispersion would produce a narrower line and thus a worse fit.
From the lower right panel in Figure \ref{multiphaseresults} we find that it
is unlikely that the clump velocity dispersion is lower than
$50$\kms. 
This is also supported by the classification tree example in Figure
\ref{fig:tree}.

As a test, we also run a small set of $100$ additional radiative
transfer simulations with random values for all parameters, except for
$\sigma_{\rm cl}$, $v_{\infty {\rm cl}}$, $P_{\rm cl}$ which are kept within the derived $1-\sigma$
constraints. \mg{Is $1\sigma$ correct?}
Consistent with the results from the random forest algorithm,
illustrated by the classification tree of Figure \ref{fig:tree}, we
find that $30$ models show $\chi^2/d.o.f.<35$.

Having constrained only 3 parameters, one might wonder why the other 11
parameters do not seem to matter. 
However, this view is misleading because merely the best-fit values of (some of) those $3$ parameters, i.e., the corresponding constraint to a given sub-region in the parameter space, make the other parameters less important.
For instance, in our case the large value of $P_{\rm cl}$ makes radiative transfer effects, and therefore other parameters such as the clump column density less relevant for the emergent line profile. However, for instance larger values of $f_{\rm cl}$ -- outside the here considered parameter range -- would re-introduce additional radiative transfer effects and thus increase the number of relevant parameters \citep[see ][for details]{2017arXiv170406278G}. 
This means that even
if for clumply models the line profile is most sensitive to only
a few parameters \citep{Gronke2016}, the rest of parameters
affect the line profile at a level that cannot be neglected,
\mg{I am not sure I understand the rest of this sentence. What do you mean there?}
i.e. we cannot randomly pick 11 random, reasonable values for these
parameters, and expect to fit the line profile while only varying the
other 3, as the results for the classification tree example in Figure
\ref{fig:tree} shows.

We cannot discard that another region of parameter space
could also yield a good fit to the observed line-profile.
This might be interesting to explore in the
future with, e.g., a higher sampling of the parameter space once new 
observations  clarify the kinematic nature of \tol \footnote{In
  Section \S \ref{sec:conclusions} we describe  in more detail what
  kind of observations could be performed in the near future to
  improve our understanding of \tol's kinematics.}.
Furthermore, considering the sparse sampling of parameter space and the results
from the random forest classifier, it is safer to consider that
the best values for $\sigma_{\rm cl}$, $v_{\infty {\rm cl}}$, $P_{\rm
  cl}$ are actually lower/upper/lower bounds, respectively.
\mg{This is interesting. Why do you think that? Could we use that to increase the quoted errors? I am worried that the referee will complain about the small errors.}

@@

\bibliographystyle{mn2e}
%\bibliography{references}
\begin{thebibliography}{}

@@

@@

@@

>>>>

\newcommand{\mg}[1]{\textcolor{blue}{\bf MG: \textit{#1}}}
\newcommand{\jefr}[1]{\textcolor{darkgreen}{\bf JEFR: \textit{#1}}}

@@

@@

We also a Kolmogorov-Smirnov (K-S) test to compare each parameter
distribution in the best $25$ models against the parent distribution
of $2500$ models. 
From a p-value $<0.05$ for a given model parameter, we conclude that
this parameter can be constrained from the observations, as the distribution for
the best $\chi^2$ models is statistically different to the
distribution from the global sample of $2500$ models.

We finally compute the best values for the constrained parameters as
the values that produce the minimum $\chi^2$.  
We estimate the 1-$\sigma$ uncertainty from a parabolic fit to the
$\chi^2$ as a function of the constrained parameters around its
corresponding minimum.

@@

@@

With $14$ free parameters our first concern is discovering which are
the parameters that can be safeley constrained by \tol's line shape.

From the K-S tests we confirm that only 3 parameters 
can be constrained by observations;
$v_{\infty,\rm{cl}}$ (p-value  $10^{-18}$), $\sigma_{\rm cl}$ (p-value
$10^{-4}$) $P_{\rm cl}$ (p-value $10^{-4}$).
This does not mean the other parameters do not influence the resulting
spectra at all; it means that they cannot be constrained from \tol's
observations.

The best models are shown in Figure \ref{multiphaseresults}.
Left column shows the difference between the integrated distributions
of the full sample (2500 input models) and the models with the lowest
$\chi^2/d.o.f.$. 
The right column shows the actual $\chi^2/d.o.f.$ 
and its corresponding parameter value. 
Under these conditions we find $\sigma_{\rm cl}=$\sigmaclump,
$v_{\infty {\rm, cl}}=$\inftyclump\ and $P_{\rm cl}=$\probaclump.
\mg{These errors seem still pretty small to me. You obtained these via a Fisher matrix of a parabolic fit, right? How good was the fit? Are the off-diagonal elements included?}
\jefr{The small uncertainty only says that the increase in
  $\chi^2/dof$ is very steep. I don't think you can get larger errors
  with the available date. By simple inspection of Figure 2 we can
  see that the errors will be small with the data we have. The
  paremeter uncertainty roughly corresponds to the width of the parabola
  when there is a difference of $\chi^2/dof\sim 1$ from the minimum. In the data we
  have there is a huge jump of $\chi^2/dof$ of $\sim 5$ between the minimum
  and the nearest data point. }

Qualitatively this result can be understood as follows.
Due to the large fraction of \lya photons being emitted within the
moving clumps ($P_{\rm cl}\sim 1$) the `intrinsic' profile represents
the kinematics of the clumps, that is, the width of the intrinsic line
is set by $\sigma_{\mathrm{cl}}$ and its median offset mainly by
$v_{\infty}$ \& $\beta_{\rm cl}$. 
Furthermore, the relatively low mean number of clumps per line of
sight ($f_{\mathrm{cl}}$) combined with the high velocity dispersion
of the clumps implies the  existence of low-density inter-clump 
regions where \lya photons can freely propagate which leads to an
emergent spectrum close to the intrinsic one as well as to the high
flux at line center as observed. 
Because the width of the observable spectrum is hence set primarily by
$\sigma_{\rm cl}$, a lower velocity dispersion would produce a
narrower line and thus a worse fit. 
From the lower right panel in Figure \ref{multiphaseresults} we find that it
is unlikely that the clump velocity dispersion is lower than
$50$\kms.

As a test, we also run a small set of $100$ additional radiative
transfer simulations with random values for all parameters, except for
$\sigma_{\rm cl}$, $v_{\infty {\rm cl}}$, $P_{\rm cl}$ which are kept within the derived $1\sigma$
constraints. We find that $30$ models show $\chi^2/d.o.f.<35$.
\mg{Is $1\sigma$ correct?}\jefr{What do you mean with
  that question?}
Consistent with the results from the random forest algorithm,
illustrated by the classification tree of Figure \ref{fig:tree},

Having constrained only 3 parameters, one might wonder why the other 11
parameters do not seem to matter. 
However, this view is misleading because merely the best-fit values of (some of) those $3$ parameters, i.e., the corresponding constraint to a given sub-region in the parameter space, make the other parameters less important.
For instance, in our case the large value of $P_{\rm cl}$ makes radiative transfer effects, and therefore other parameters such as the clump column density less relevant for the emergent line profile. However, for instance larger values of $f_{\rm cl}$ -- outside the here considered parameter range -- would re-introduce additional radiative transfer effects and thus increase the number of relevant parameters \citep[see ][for details]{2017arXiv170406278G}. 
This means that even
if for clumply models the line profile is most sensitive to only
a few parameters \citep{Gronke2016}, the rest of parameters
affect the line profile at a level that cannot be neglected,
\mg{I am not sure I understand the rest of this sentence. What do you mean there?}
i.e. we cannot randomly pick 11 random values for these
parameters and expect to fit the line profile while only varying the
other 3

We cannot discard that another region of parameter space
could also yield a good fit to the observed line-profile.
This might be interesting to explore in the
future with, e.g., a higher sampling of the parameter space once new 
observations  clarify the kinematic nature of \tol \footnote{In
  Section \S \ref{sec:conclusions} we describe  in more detail what
  kind of observations could be performed in the near future to
  improve our understanding of \tol's kinematics.}.

@@

\bibliographystyle{mn2e}
%\bibliography{references}

\begin{thebibliography}{}

@@

\bibitem[\protect\citeauthoryear{{Gronke}, {Dijkstra}, {McCourt} \&
  {Oh}}{{Gronke} et~al.}{2017}]{2017arXiv170406278G}
{Gronke} M.,  {Dijkstra} M.,  {McCourt} M.,    {Oh} S.~P.,  2017, ArXiv
  e-prints

@@

\bibitem[\protect\citeauthoryear{{McCourt}, {Oh}, {O'Leary} \&
  {Madigan}}{{McCourt} et~al.}{2016}]{2016arXiv161001164M}
{McCourt} M.,  {Oh} S.~P.,  {O'Leary} R.~M.,    {Madigan} A.-M.,  2016, ArXiv
  e-prints

\bibitem[\protect\citeauthoryear{{McKee} \& {Ostriker}}{{McKee} \&
  {Ostriker}}{1977}]{1977ApJ...218..148M}
{McKee} C.~F.,  {Ostriker} J.~P.,  1977, \apj, 218, 148

@@

\appendix
\section{Random Forest Classification}
\label{appendix}

As a complement to the K-S tests on the multiphase data 
we apply a random forest classification algorithm to find the
more relevant parameters in the model to produce a low $\chi^2$
result.

Using the total number of degrees of freedom, $d.o.f.=104$ we divide
the results in two classes: low $\chi^2/d.o.f.<35$ and high
$\chi^2/d.o.f.>35$. \mg{Why was $35$ chosen as a boundary?}\jefr{We
  chose $35$ to match the cut that we used to make the  integrated
  probability distributions. However the nuber itself is
  arbitrary. Ideally we should only chose models with $\chi^/dof$
  higher by $2$ than the lowest $\chi^2/dof$, but we don't have any
  models in that range.}
The algorithm uses $500$ trees for the classification and a
maximum of $3$ depth levels. 
To check for stability we repeat the computation $10$ times by
randomly subsampling the input data to use $80\%$ of the data as
the training set.

Figure \ref{fig:tree} shows as an illustration the results for a
single classification tree.
The tree starts with $28$ and $1962$ models in the low and high
$\chi^2/d.o.f.$ classes, respectively. 
We find that the best classification
yields $13$ and $44$ models in the low and high $\chi^2/d.o.f.$
classes, respectively, after selecting for $v_{\infty,{\rm cl}}<157.0$ \kms,
$\sigma_{{\rm cl}}>55.6$ \kms and $P_{\rm cl}>0.683$. 
The split in two classes is not perfect. 
This means that these 3 parameters are the most useful parameters
in finding models with low $\chi^2/d.o.f.$.
Not having a perfect classification using these three parameter only,
shows that some combinations of the other 11 parameters are also
relevant to find a good fit.

The results of the random forest classifier yield that the clump outflow velocity
$v_{\infty,\rm{cl}}$, the clump velocity dispersion $\sigma_{\rm cl}$ and the probability
that the \lya emission comes from the clumps $P_{\rm cl}$, are the
most influential parameters in finding a model with low
$\chi^2/d.o.f.$.

Furthermore, considering the sparse sampling of parameter space and the results
from the random forest classifier, it is safer to consider that
the best values for $\sigma_{\rm cl}$, $v_{\infty {\rm cl}}$, $P_{\rm
  cl}$ are actually lower/upper/lower bounds, respectively.
\mg{This is interesting. Why do you think that? Could we use that to
  increase the quoted errors? I am worried that the referee will
  complain about the small errors.} \jefr{The referee didn't complain
  about the small errors.}

\begin{figure*}
\begin{center}
\includegraphics[width=0.95\textwidth]{tree_result.pdf}
\caption{Classification tree example. 
The aim is to find the parameters that can be used to separate the
results in two classes: low $\chi^2/d.o.f. < 35$ and high $\chi^2/d.o.f>35$. 
Each box describes the condition over the parameter of interest, the
level of sample impurity, the total number of samples and the value of the
number of samples in each class.  
In this example we randomly sample $80\%$ of the full data set of models to
start with $28$ models in the first class and
$1962$ models in the second class; the best way to increase the
probability to find a result with low $\chi^2/d.o.f.$ ($13$ models,
fourth bottom box, from left to the right)
is having the clump radial velocity, $v_{\infty,{\rm cl}}<157.0$ \kms;
the clump velocity dispersion, $\sigma_{{\rm cl}}>55.6$ \kms, and the
probability to have a \lya photon emitted in a clump, $P_{\rm cl}>0.683$.
The final classification is not perfect, showing that all other
parameters are also important but these three are the most relevant.  
\label{fig:tree}}
\end{center}
\end{figure*}
